
### Project background
Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a 
large amount of data about personal activity relatively inexpensively. 
These type of devices are part of the quantified self movement - a group of 
enthusiasts who take measurements about themselves regularly to improve their health.  

Goal of the project is to use data from accelerometers on the belt, forearm, arm, and 
dumbell of 6 participants and quantify how well they do it (manner in which they did the exercise)

### Data source 
The data for this project come from this source: http://groupware.les.inf.puc-rio.br/har

The training data for this project are available here: https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv
The test data are available here: https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv

### Getting data
The training and test files are downloaded and saved in the working directory

```{r, message=FALSE,warning=FALSE}
library(caret)
library(randomForest)
setwd("d:/readplease/courseradatascience/machinelearning")
```

```{r}
intrain <- read.csv('pml-training.csv', header = TRUE, sep = ",", na.strings = c("NA", "#DIV/0!"))
intest <- read.csv('pml-testing.csv', header = TRUE, sep = ",", na.strings = c("NA", "#DIV/0!"))
dim(intrain)
dim(intest)
```

### Cleaning data
The data files have 160 columns. I have used the column names and selected only the ones which are measurement related. The columns with almost all values as NA are removed.  

```{r}
set.seed(123)
sensorvalues = grep(pattern = "arm|belt|dumbbell|forearm|classe", names(intrain))
train = intrain[, c(sensorvalues)]
missing = is.na(train)
unwanted = which(colSums(missing) > 10000)
training = train[, -unwanted]
dim(training)
```

Now the data file has 53 columns. These columns are used to predict the classe.

### Splitting the training file into train and test
The training file is split into train (70%)and test (30%)

```{r}
split = createDataPartition(training$classe, p = 0.7,list=FALSE) 
training.set = training[ split,]
testing.set = training[ -split,]
```

### Using randomForest algorithm to predict
randomForest is used to train a model and the model is used to predict the results
```{r}
rf = randomForest(classe~., data=training.set, ntree = 500)
prediction = predict(rf, newdata = testing.set)
```

```{r}
confusionMatrix(prediction, testing.set$classe)
```
The accuracy of the model is 99.44%

### Using the model to predict the classe variable for test file

```{r}
answers = predict(rf, newdata = intest)
answers
```

The predictions for the 20 record in the test file is given above.